Taking Dyads Seriously

### 2022-08-03 paper/toSubmit/netsMatter.tex
cleanup

% references to graphics
\makeatletter
% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}

@@

% Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another and not passing any auxiliary information to the model about how observations may be interdependent. For example, in this design, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States. Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan. While most scholars begin with the assumption that each dyadic interaction takes place in isolation of the others, we know this assumption to be false both in theory and in practice. Relational data comes with an explicit structure that generally leads to particular types of dependencies. The importance of accounting for the underlying structure of our data has been a lesson well understood, at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}. As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.

@@

\begin{figure}
	\begin{tabular}{cc}
		\textsc{Sender Heterogeneity} & \textsc{Receiver Heterogeneity} \\
		\includegraphics[width=.4\textwidth]{graphics/adjRowDep.png} &
		\includegraphics[width=.4\textwidth]{graphics/adjColDep.png} \\
		\textsc{Sender-Receiver Covariance} & \textsc{Reciprocity} \\
		\includegraphics[width=.4\textwidth]{graphics/adjRowColCovar.png} &
		\includegraphics[width=.4\textwidth]{graphics/adjRecip.png} \\
	\end{tabular}
	\caption{Nodal and dyadic dependencies in relational data.}
	\label{fig:adjMatDeps}
\end{figure}

@@

\begin{figure}[ht]
	\includegraphics[width=.6\textwidth]{graphics/stochEquiv_v2.pdf}
	\caption{Higher-order dependence patterns in a network.}
	% need a simpler caption here
	\label{fig:thirdDeps}
\end{figure}

@@

% International Trade, we might observe homophily wherein states with open economies and an absence of Non-Tariff Barriers are more likely to trade with each-other all else equal; alternatively we might observe heterophily in the sense that states that have greater endowments of capital are more likely to trade with states with greater endowments of land or labor.

%  We would also likely observe stochastic equivalence in an interstate war network wherein countries who are in conflict with the same target country are unlikely to fight one another. In this case, countries with (formal or informal) alliances are a member of the same group due to similar conflictual ties to other actors in the system.

@@

\begin{figure}
	\centering
	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
	\label{fig:ameBias}
	\includegraphics[width=1\textwidth]{graphics/figure3.pdf} \\
\end{figure}

@@

\begin{figure}
	\centering
	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib}
	\includegraphics[width=1\textwidth]{graphics/figure4.pdf} \\
\end{figure}

@@

\begin{figure}
	\centering
	\caption{Distribution of correlation between missing variable and multiplicative random effect in AME across the 1,000 simulations. Vertical line through the distribution represents the median value across the simulations.}
	\label{fig:ameCorr}
	\includegraphics[width=1\textwidth]{graphics/figure5.pdf} \\
\end{figure}

@@

\begin{figure}
	\centering
	\begin{tabular}{cc}
		\multicolumn{2}{l}{\textbf{\tiny{Reiter \& Stam (2003)}}} \\
		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_pr_outSample.pdf} \\
		\multicolumn{2}{l}{\textbf{\tiny{Weeks (2012)}}} \\
		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_pr_outSample.pdf} \\
		\multicolumn{2}{l}{\textbf{\tiny{Gibler (2017)}}} \\
		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_pr_outSample.pdf} \\
	\end{tabular}
	\caption{Assessments of out-of-sample predictive performance using ROC curves, PR curves, and separation plots.}
	\label{fig:perf}
\end{figure}

@@

\begin{figure}[ht]
	\begin{tabular}{cc}
	\includegraphics[width=.7\textwidth]{graphics/figure7a.pdf} &
	\includegraphics[width=.3\textwidth]{graphics/figure7b.pdf} \\
	\end{tabular}
	\caption{Estimates of sender random effects ($a_{i}$) from AME for the Reiter \& Stam (2003) model. Positive values indicate that the particular country is more likely to be involved in conflict than predicted by the covariates in the model. Negative values indicate that the country is less likely to be involved in conflict. The left-panel shows $a_{i}$ estimates for all countries and right-panel highlights the top ten countries in terms of positive and values of $a_{i}$.}
	\label{fig:reiter_stam_aEff}
\end{figure}
\FloatBarrier

@@

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{graphics/figure8.pdf}
	\caption{ Visualization of multiplicative effects for Weeks (2012). Each circle designates a country and the color corresponds to the legend at the center of the visualization. Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries. The inner ring clusters countries by the similarity of who they receive conflict from.
	}
	\label{fig:weekscirc}
\end{figure}
\FloatBarrier

@@

\begin{figure}
	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
	\includegraphics[width=\textwidth]{graphics/figure9.pdf}
 	\label{fig:gibmargeff}
 \end{figure}

@@

\input{graphics/appendix_tableB1}
\FloatBarrier

@@

\input{graphics/appendix_tableB2}
\FloatBarrier

@@

\input{graphics/appendix_tableB3}
\FloatBarrier

@@

\begin{figure}[ht]
	\centering
	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
	\label{fig:ameBias_asa}
	\includegraphics[width=1\textwidth]{graphics/appendix_figureC1.pdf} \\
\end{figure}

\begin{figure}[ht]
	\centering
	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib_asa}
	\includegraphics[width=1\textwidth]{graphics/appendix_figureC2.pdf} \\
\end{figure}

@@

\begin{figure}[ht]
\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.4 in this case.}
\label{fig:ameBias_corrMed}
\includegraphics[width=1\textwidth]{graphics/appendix_figureD1a.pdf} \\
\end{figure}

\begin{figure}[ht]
\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.7 in this case.}
\label{fig:ameBias_corrHi}
\includegraphics[width=1\textwidth]{graphics/appendix_figureD1b.pdf} \\
\end{figure}

>>>>

% references to graphics
\makeatletter
% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/}}
% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/}}

@@

@@

\begin{figure}
	\begin{tabular}{cc}
		\textsc{Sender Heterogeneity} & \textsc{Receiver Heterogeneity} \\
		\includegraphics[width=.4\textwidth]{adjRowDep.png} &
		\includegraphics[width=.4\textwidth]{adjColDep.png} \\
		\textsc{Sender-Receiver Covariance} & \textsc{Reciprocity} \\
		\includegraphics[width=.4\textwidth]{adjRowColCovar.png} &
		\includegraphics[width=.4\textwidth]{adjRecip.png} \\
	\end{tabular}
	\caption{Nodal and dyadic dependencies in relational data.}
	\label{fig:adjMatDeps}
\end{figure}

@@

\begin{figure}[ht]
	\includegraphics[width=.6\textwidth]{stochEquiv_v2.pdf}
	\caption{Higher-order dependence patterns in a network.}
	% need a simpler caption here
	\label{fig:thirdDeps}
\end{figure}

@@

@@

\begin{figure}
	\centering
	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
	\label{fig:ameBias}
	\includegraphics[width=1\textwidth]{figure3.pdf} \\
\end{figure}

@@

\begin{figure}
	\centering
	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib}
	\includegraphics[width=1\textwidth]{figure4.pdf} \\
\end{figure}

@@

\begin{figure}
	\centering
	\caption{Distribution of correlation between missing variable and multiplicative random effect in AME across the 1,000 simulations. Vertical line through the distribution represents the median value across the simulations.}
	\label{fig:ameCorr}
	\includegraphics[width=1\textwidth]{figure5.pdf} \\
\end{figure}

@@

\begin{figure}
	\centering
	\begin{tabular}{cc}
		\multicolumn{2}{l}{\textbf{\tiny{Reiter \& Stam (2003)}}} \\
		\includegraphics[width=.4\textwidth]{figure6_reiter_stam_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{figure6_reiter_stam_pr_outSample.pdf} \\
		\multicolumn{2}{l}{\textbf{\tiny{Weeks (2012)}}} \\
		\includegraphics[width=.4\textwidth]{figure6_weeks_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{figure6_weeks_pr_outSample.pdf} \\
		\multicolumn{2}{l}{\textbf{\tiny{Gibler (2017)}}} \\
		\includegraphics[width=.4\textwidth]{figure6_gibler_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{figure6_gibler_pr_outSample.pdf} \\
	\end{tabular}
	\caption{Assessments of out-of-sample predictive performance using ROC curves, PR curves, and separation plots.}
	\label{fig:perf}
\end{figure}

@@

\begin{figure}[ht]
	\begin{tabular}{cc}
	\includegraphics[width=.7\textwidth]{figure7a.pdf} &
	\includegraphics[width=.3\textwidth]{figure7b.pdf} \\
	\end{tabular}
	\caption{Estimates of sender random effects ($a_{i}$) from AME for the Reiter \& Stam (2003) model. Positive values indicate that the particular country is more likely to be involved in conflict than predicted by the covariates in the model. Negative values indicate that the country is less likely to be involved in conflict. The left-panel shows $a_{i}$ estimates for all countries and right-panel highlights the top ten countries in terms of positive and values of $a_{i}$.}
	\label{fig:reiter_stam_aEff}
\end{figure}
\FloatBarrier

@@

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{figure8.pdf}
	\caption{ Visualization of multiplicative effects for Weeks (2012). Each circle designates a country and the color corresponds to the legend at the center of the visualization. Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries. The inner ring clusters countries by the similarity of who they receive conflict from.
	}
	\label{fig:weekscirc}
\end{figure}
\FloatBarrier

@@

\begin{figure}
	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
	\includegraphics[width=\textwidth]{figure9.pdf}
 	\label{fig:gibmargeff}
 \end{figure}

@@

\input{appendix_tableB1}
\FloatBarrier

@@

\input{appendix_tableB2}
\FloatBarrier

@@

\input{appendix_tableB3}
\FloatBarrier

@@

\begin{figure}[ht]
	\centering
	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
	\label{fig:ameBias_asa}
	\includegraphics[width=1\textwidth]{appendix_figureC1.pdf} \\
\end{figure}

\begin{figure}[ht]
	\centering
	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib_asa}
	\includegraphics[width=1\textwidth]{appendix_figureC2.pdf} \\
\end{figure}

@@

\begin{figure}[ht]
\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.4 in this case.}
\label{fig:ameBias_corrMed}
\includegraphics[width=1\textwidth]{appendix_figureD1a.pdf} \\
\end{figure}

\begin{figure}[ht]
\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.7 in this case.}
\label{fig:ameBias_corrHi}
\includegraphics[width=1\textwidth]{appendix_figureD1b.pdf} \\
\end{figure}
